Figures and Analysis for 2023 Manuscript

### 2023-12-01 data-raw/Manuscript2023_Workflow.Rmd
Added the alternative (new) figures to vignette and fixed lack of exclusion criteria for plot method

other_coefs <- read_csv(
  "C:/Users/GPADILLA/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/Figures/CvTdb_fullCoefs_20230906.csv")

other_coefs %>% dplyr::select(model, method, Chemical, Species)

write_csv(my_coefs, "C:/Users/GPADILLA/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/Figures/CvTdb_fullCoefsAIC_2023.10.30.csv")

@@

@@

plot_grid(pl5A, pl5B, labels = c("A", "B"),
                     align = "h", axis = "bt", rel_widths = c(1, 1.5))

figure5 <- plot_grid(pl5A, pl5B, labels = c("A", "B"),
                     align = "h", axis = "bt", rel_widths = c(1, 1.5))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "Figure5_onlyDetects_2023-10-02.png"),
       plot = figure5,
       bg = "white",
       height = 3.25,
       width = 6.5,
       units = "in",
       dpi = 300)

@@

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "PredictedObserved_compartmentModels_2023-10-02.png"),
       height = 6, width = 6)

@@

```{r fig5Alt}
this_preds <- inner_join(predict(my_pk), winmodel) %>%
  rowwise() %>% mutate(foldPred = Conc_est/Conc) %>%
  distinct(model, method, Chemical, Species, Route, Media, Dose, Time, foldPred) %>%
  group_by(model, method, Chemical, Species, Route, Media, Dose, Time) %>%
  summarize(avgFE = mean(foldPred, na.rm = TRUE))
this_preds %>%
  inner_join(combined_summary, by = join_by(Chemical, Species,
                                            Route, Media,
                                            Dose, Time)) %>%
  filter(!is.na(model), avgFE > 0) %>% 
  # For percentages
  # ungroup() %>%
  # filter(model %in% c("model_flat")) %>%
  # count((foldPred >= 0.5 & foldPred <= 2) & !(foldConc >= 0.5 & foldConc <= 2)) %>%
  # {100*.[[2,2]]/(.[[1,2]] + .[[2,2]])}
  # For Plotting
  ggplot(aes(
    y = avgFE,
    x = twofold_95
  )) +
  geom_bin2d(bins = 40, color = NA) +
  geom_hline(yintercept = c(0.5, 2), linetype = "dashed") +
  scale_y_log10(labels = scales::label_math(format = log10),
                limits = c(0.0001, 10000)) +
  scale_fill_viridis_c(option = "cividis",
                       limits = c(1, 200),
                       oob = scales::oob_squish) +
  labs(y = "Model Error",
       x = "95% of data within 2-fold") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

@@

@@

my_coefs %>%
  mutate(model = 
           case_when(
             model == "model_1comp" ~ "1-compartment",
             model == "model_2comp" ~ "2-compartment",
             model == "model_flat"  ~ "Null"
           )) %>%
  ggplot(aes(x = sigma_value/n_obs)) +
  geom_histogram(bins = 15,
                 color = NA, fill = "grey5") +
  scale_x_log10(labels = label_log()) +
  facet_grid(cols = vars(model)) +
  labs(x = "Size-normalized Sigma Value",
       y = "Number of Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank())

@@

```{r sfig6-examples-of-fits}
pl <- plot(my_pk, use_scale_conc = TRUE, n_interp = 10)
names(pl) # Chemical Species observations oservation_plot predicted predicted_plot final_plot

s6_plA <- pl %>% filter(Chemical %in% "DTXSID90382620") %>%
  pull(observation_plot) %>% .[[1]]
s6_plB_mouse <- pl %>% filter(Chemical %in% "DTXSID4020533",
                              Species %in% "mouse") %>% pull(final_plot) %>% .[[1]]
s6_plB_rat <- pl %>% filter(Chemical %in% "DTXSID4020533",
                              Species %in% "rat") %>% pull(final_plot) %>% .[[1]]
s6_plC <- pl %>% filter(Chemical %in% "DTXSID8021359") %>%
  pull(final_plot) %>% .[[1]]
s6_plD <- pl %>% filter(Chemical %in% "DTXSID8025337") %>%
  pull(final_plot) %>% .[[1]]

# Set colors palletes for each
# panel A : Oranges
panelA_cols <- brewer.pal(8, "Oranges")[3:8]
# panel B : Greens
panelB_cols <- brewer.pal(9, "Greens")[2:9]
# panel C : Blue (only 2 values) 
# panel D : Purple (6 values)
panelD_cols <- brewer.pal(9, "Purples")[4:9]

s6_plA <- s6_plA + 
  labs(title = "fluorotelomer acid 7+3") +
  scale_color_manual(values = panelA_cols, aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

@@

# Am making this 2X the size I need it to be 
# so the points are a reasonable size when I scale it down
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ExampleFitsPlots.png"),
       width = 13,
       height = 14,
       units = "in")

>>>>

@@

table(combined_summary$twofold_95)

pl5A <- combined_summary %>% ungroup() %>%
  count(twofold_95) %>%
  ggplot(aes(x = twofold_95, fill = twofold_95, y = n)) +
  geom_col(position = position_stack(),
           color = "black", linewidth = 1) +
  coord_cartesian(ylim = c(0, 4000)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = c(0, 1000, 2000, 3000),
                     labels = c("0", "1", "2", "3")) +
  scale_fill_manual(values = c("black", "white")) +
  theme_classic() +
  labs(x = "95% of observations\nwithin factor of 2",
                y = "Observations\n(thousands)") +
  theme(text = element_text(size = 10),
        panel.border = element_rect(color = "black", linewidth = 1, fill = NA),
        panel.background = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.4),
        axis.title = element_text(face = "bold"),
        axis.line = element_blank(),
        plot.background = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")

@@

plot_grid(pl5A, pl5B, labels = c("A", "B"),
                     align = "h", axis = "bt", rel_widths = c(1, 2))

figure5 <- plot_grid(pl5A, pl5B, labels = c("A", "B"),
                     align = "h", axis = "bt", rel_widths = c(1, 1.75))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "Figure5alt_onlyDetects_2023-10-02.png"),
       plot = figure5,
       bg = "white",
       height = 3.25,
       width = 6.5,
       units = "in",
       dpi = 300)

@@

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "PredictedObserved_compartmentModels_2023-11-30.png"),
       height = 6, width = 6)

@@

```{r fig5Alt}
# Redo combined summary
combined_summary <- bind_rows(sud_summary, iad_summary) %>%
  rowwise() %>%
  mutate(twofold_95 = ifelse((conc_avg_calc + (2*conc_sd_calc))/conc_avg_calc < 2,
                             TRUE, FALSE))

# Now
this_preds <- inner_join(predict(my_pk), winmodel) %>%
  rowwise() %>% mutate(foldPred = Conc_est/Conc) %>%
  distinct(model, method, Chemical, Species, Route, Media, Dose, Time, foldPred) %>%
  group_by(model, method, Chemical, Species, Route, Media, Dose, Time) %>%
  summarize(avgFE = mean(foldPred, na.rm = TRUE))

pl_7Alt <- this_preds %>%
  inner_join(combined_summary, by = join_by(Chemical, Species,
                                            Route, Media,
                                            Dose, Time)) %>%
  filter(!is.na(model), avgFE > 0) %>% 
  # 
  # For Plotting
  ggplot(aes(
    y = avgFE,
    x = twofold_95
  )) +
  geom_bin2d(bins = 40, color = NA) +
  geom_hline(yintercept = c(0.5, 2), linetype = "dashed", linewidth = 1, color = "red3") +
  scale_y_log10(labels = scales::label_math(format = log10),
                limits = c(0.0001, 10000)) +
  scale_fill_viridis_c(option = "cividis",
                       limits = c(1, 200),
                       oob = scales::oob_squish) +
  labs(y = "Model Error",
       x = "95% of data within 2-fold") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

@@

this_preds %>%
  inner_join(combined_summary, by = join_by(Chemical, Species,
                                            Route, Media,
                                            Dose, Time)) %>%
  filter(!is.na(model), avgFE > 0) %>% ungroup() %>%
  filter(model == "model_2comp") %>%
  count(!(avgFE >= 0.5 & avgFE <= 2) & !twofold_95) %>%
  {100*.[[2,2]]/(.[[1,2]] + .[[2,2]])} %>% round(digits = 1)

# Table

my_table <- data.frame(
  c("Reasonable model error and observed variability",
    "Data variability high, model error reasonable",
    "Model error high, data variation reasonable",
    "Both model error and data variability are high"),
  c(37.2, 6.5,  41.4, 14.8),
  c(53.6, 3.5,  32.4, 10.4),
  c(11.0, 16.4, 46.6, 26.0),
  c(44.2, 5.4,  37.4, 13.0))

dual_plot <- plot_grid(pl_7Alt, table_plot, ncol = 1, align = "v",
          rel_heights = c(1, 0.5))

dual_plot
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ModelPerformance_vDataVariability_wTable_2023-11-30.png"),
       plot = dual_plot,
       height = 4.5,
       width = 6.5,
       device = "png",
       bg = "white",
       dpi = 300,
       units = "in")

@@

my_coefs %>%
  mutate(model = 
           case_when(
             model == "model_1comp" ~ "1-compartment",
             model == "model_2comp" ~ "2-compartment",
             model == "model_flat"  ~ "Null"
           )) %>%
  ggplot(aes(x = sigma_value/n_obs)) +
  geom_histogram(bins = 15,
                 color = NA, fill = "grey5") +
  scale_x_log10(labels = scales::label_math(format = log10)) +
  facet_grid(cols = vars(model)) +
  labs(x = "Size-normalized Sigma Value",
       y = "Number of Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank())

@@

```{r sfig6-examples-of-fits}
pl <- plot(my_pk, use_scale_conc = TRUE, n_interp = 10)
names(pl) # Chemical Species observations observation_plot predicted predicted_plot final_plot

s6_plA <- pl %>% filter(Chemical %in% "DTXSID3031860") %>%
  pull(observation_plot) %>% .[[1]]
s6_plB_mouse <- pl %>% filter(Chemical %in% "DTXSID4020533",
                              Species %in% "mouse") %>% pull(final_plot) %>% .[[1]]
s6_plB_rat <- pl %>% filter(Chemical %in% "DTXSID4020533",
                              Species %in% "rat") %>% pull(final_plot) %>% .[[1]]
s6_plC <- pl %>% filter(Chemical %in% "DTXSID8021359") %>%
  pull(final_plot) %>% .[[1]]
s6_plD <- pl %>% filter(Chemical %in% "DTXSID8025337") %>%
  pull(final_plot) %>% .[[1]]

# Set colors palettes for each
# panel A : Oranges
panelA_cols <- brewer.pal(8, "Oranges")[3:8]
# panel B : Greens
panelB_cols <- brewer.pal(9, "Greens")[2:9]
# panel C : Blue (only 2 values) 
# panel D : Purple (6 values)
panelD_cols <- brewer.pal(9, "Purples")[4:9]

s6_plA <- s6_plA + 
  labs(title = "perfluorodecanoic acid") +
  scale_color_manual(values = panelA_cols, aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

@@

# Am making this 2X the size I need it to be 
# so the points are a reasonable size when I scale it down
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ExampleFitsPlots_11.30.2023.png"),
       width = 13,
       height = 14,
       units = "in")
